 

### 2019-07-26 Manuscript/Manuscript R1/Manuscript.Rmd
Update old version

figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

header-includes:
  - \DeclareRobustCommand{\VANDER}[3]{#2} 
  - \usepackage{multirow}
  - \usepackage{rotating}
  - \usepackage{color}
  - \usepackage{subcaption}
   #% set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)

@@

```{r, echo = FALSE}
library(papaja)
library(circular)
library(bpnreg)
library(haven)
library(tikzDevice)
library(plotrix)
library(tidyr)
library(dplyr) #sample_n()
library(MASS) #mvrnorm()
library(kableExtra) #latex tables
library(plotrix)
library(caret) #cross-validation
library(coda) #geweke diagnostic
```

\section{Introduction}\label{Introduction}

In social sciences the use of cylindrical data is very common. Such data consist
of a linear and a circular component. @gurtman2011reasoning refers to such data
as vectors, with a directional measure (i.e., the circular component) and a
measure indicating the magnitude (i.e., linear component). Many established
models in psychology are often referred to as circular or circumplex models, but
those models are cylindrical. Examples of such cylindrical models are the
interpersonal circle/circumplex
[@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
affect [@russell1980circumplex], the circumplex of human emotion
[@plutchik1997general] or the model of human values [@schwartz1992values].
\newline
\indent Also, many of the more recent types of data that are studied in psychology are
cylindrical. For example, research on human navigation uses data where distance
(i.e., linear component) and direction (i.e., the linear component) are of
interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
consist of both the direction (i.e., the circular variable) and the duration
(i.e., the linear variable)(e.g., @rayner200935th). Apart from the social
sciences data with a circular and linear outcome more commonly occur in
meteorology [@garcia2013exploring], ecology [@garcia2014test] or
marine research [@lagona2015hidden]
\newline
\indent Up until now researchers studying cylindrical data had to rely on linear
statistical methods to analyze their research results. However, lately more and more
of these researchers acknowledge that linear methods are not sufficient and call
for new methods [@gurtman2011reasoning; @pennings2017complexity;
@wright2009integrating] that take into account both the circular and the linear
component of these data.
\newline
\indent The aim in the present paper is twofold. Firstly, we we intend to fill
the above mentioned gap in the literature by showing the use of cylindrical
models can benefit the analysis of circumplex data and cylindrical data in
psychology in general. More specifically we will show these benefits for
interpersonal teacher data from the field of educational psychology. Apart from
modelling the dependence between the linear and circular component of a
cylindrical variable we would also like to predict the two components from a set
of covariates in a regression model. Our second aim therefore is to adapt
several existing cylindrical models in such a way that they include a regression
of both the linear and circular component of a cylindrical variable onto a set
of covariates. From now on we will therefore refer to the components of the
cylindrical variable as outcome components. These adapted circular models are
then used  to analyse the teacher data.

\subsection{Cylindrical data}

Data that consist of a variable and a circular variable is called
cylindrical data. A circular variable is different from the linear variable in
the sense that it is measured on a different scale. Figure \ref{circline} shows
the difference between a circular scale (right) and a linear scale (left). The
most important difference is that on a circular scale the datapoints 0\(^\circ\)
and 360\(^\circ\) are connected and in fact represent the same number while on a
linear scale the two ends, \(-\infty\) and \(\infty\), are not connected and
consequently the values 0\(^\circ\) and 360\(^\circ\) are located on different
places on the scale. Both circular data and cylindrical data require special
analysis methods due to this periodicity in the scale of a circular variable
(see e.g. @fisher1995statistical for an introduction to circular data and
@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
elaborate overview).\newline
\indent In the literature, several methods have been put forward to model the
relation between the linear and circular component of a cylindrical variable.
Some of these are based on regressing the linear component onto the circular
component using the following type of relation: \[y = \beta_0 +
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
linear component and \(\theta\) the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
different way, e.g. by specifying a multivariate model for several linear and
circular variables and modelling their covariance matrix
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
example, @abe2017tractable introduce a cylindrical distribution based on a
Weibull distribution for the linear component and a sine-skewed von Mises
distribution for the circular component and link these through their respective
shape and concentration parameters. However, none of the methods that have been
proposed thus far include additional covariates onto which both the circular and
linear component are regressed.

@@

\section{Four Cylindrical Regression Models}\label{Models}

One of the goals of this paper is to show the benefits of cylindrical methods
for the analysis of circumplex data and cylindrical data in psychology in
general. To do so we decide to focus on four cylindrical models. The models were
selected for their relatively low complexity and the ease with which a
regression structure could be incorporated. But also because they show different
ways of modelling the the linear and circular outcome and thereby illustrate a
wider range of cylindrical models available in the literature. The
cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_q$ for the linear and
circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
based on a construction by @mastrantonio2015bayesian, while the other models are
extensions of the models from @abe2017tractable and @mastrantonio2018joint. The
four cylindrical  models are introduced separately in the subsections below.
However, to provide a more succinct overview and comparison of the four models, Table
\ref{TableModels} gives an overview of the similarities and differences between
the models.

\begin{sidewaystable}[h]
\centering
\caption{Comparison of the four cylindrical regression models} 
\begin{tabular}{lllll}
  \noalign{\smallskip}\hline\noalign{\smallskip}
\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
$\Theta$ & &&&\\
$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
$\:\:$Shape & symmetric, & assymetric, & assymetric, & assymetric, \\
            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
$Y$& &&&\\
$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(0, + \infty)$\\
$\:\:$Shape & symmetric, & symmetric, & assymetric, & assymetric, \\
            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
$\Theta$-$Y$ dependence &                                   &                                   & & \\
                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\

@@

& \xi\\ \xi & 1 \end{bmatrix}\), allowing for multimodality and
assymetry/skewness.\newline

@@

\begin{figure}[]
  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to. The square indicates the inflection point of the regression live.}
  \label{circregline}
\end{figure}

\subsection{The Modified Abe-Ley Model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. It concerns a combination
of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
model defined only on the positive real half-line \([0, + \infty)\) and thus can
not be negative.\newline
\indent In this model we predict the linear scale parameter and circular
location parameter, both of which we can express in terms of
covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\gamma_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\).\newline
\indent For the teacher data, the predicted values for the circular outcome in
the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
The conditional distribution for the linear outcome is Weibull, meaning that we
can use methods from survival analsis to interpret the effect of a predictor. In
survival analysis a 'survival' function is used in which time is plotted against
the probability of survival of subjects suffering from a specific medical
condition. In the teacher data we can thus compute the probability of a teacher
having a specific strength on the IPC. This probability is computed using the
'survival-function' defined as  \[\exp(-\alpha
y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
function we also see that the circular concentration parameter \(\kappa\) and
the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
and strength of interpersonal behavior through the parameters $\beta_{0}$ and
$\gamma_0$ as well as the effect of self-efficacy on type and strength of
teacher behavior through the parameters, $\beta_{1}$ and $\gamma_1$.\newline
\indent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. The model contains \(m\)
  independent circular outcomes and \(w\) independent linear outcomes. The
  circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,
  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Thus also in this model the linear outcome is
  defined constrained to be positive. Because the GPN distribution is modelled
  using a so-called augmented representation (see also the description of the
  CL-PN and CL-GPN models) it is convenient to use a similar tactic for
  modelling the multivariate skew normal distribution. As in
@mastrantonio2018joint dependence between the linear and circular outcome is
  created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
  and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
  distribution.\newline
\indent This means that we have a shared mean vector and variance-covariance
matrix for the linear and circular outcome(s), much like having multiple
outcomes in a MANOVA (multivariate analysis of variance) model. In our
regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
covariates. The mean vector then becomes \(\boldsymbol{M}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and
$\boldsymbol{x}_i$ is a $g + 1$ dimensional vector containing the value 1 to
estimate an intercept and the $g$ covariate values. This means that in contrast
to the other three models, we have to use the same set of predictors for the
circular and linear outcome.\newline
\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix}
\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} &
\beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\). The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integretion or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
linear outcomes in the GPN-SSN model are
\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
\beta_{1_{s^{I}}}\text{SE}_i),\]

@@

\noindent The GPN-SSN model thus allows us to assess the average type and
strength of interpersonal behavior through the parameters $\beta_{0_{s^{I}}}$,
$\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
on type and strength of teacher behavior through the parameters,
$\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
because the type and strength of interpersonal behavior are modelled together
using a multivariate normal distribution we can through its variance-covariance
matrix also assess the dependence between the type and strength of interpersonal
behavior. \newline
\indent We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

\subsection{Model Fit Criterion}\label{Modelfit}

For the four cylindrical models we focus on their out-of-sample predictive
performance to determine the fit of the model. To do so we use k-fold
cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
of the sample) is used once as a holdout set and 9 times as part of a training
set. The analysis will thus be performed 10 times, each time on a different
training set.\newline
\indent A proper criterion to compare out-of-sample predictive performance is
the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
value of this criterion, the better the predictive performance of the model.
Using ML estimates this criterion can be computed as follows:
\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
\end{equation}
\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
Using posterior samples the criterion is similar to the log pointwise predictive
density (lppd) [@BDA, p. 169] and can be computed as:
\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
\end{equation}
\noindent where \(B\) is the amount of posterior samples and
\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
parameters for the \(j^{th}\) iteration. Because the joint density and thus also
the likelihood for the modified GPN-SSN model is not
available in closed form [@mastrantonio2018joint] we compute the PLSL for the
circular and linear outcome separately for all models. Note that although we fit
the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
prior information into account when assessing model fit with the PLSL. According
to @BDA this is not necessary since we are assessing the fit of a model
to data, the holdout set, only. They argue that the prior in such case is only
of interest for estimating the parameters of the model but not for determining
the predictive accuracy.\newline
\indent For each of the four cylindrical models and for each of the 10
cross-validation analyses we can then compute a PLSL for the circular and linear
outcome by using the conditional log-likelihoods of the respective outcome (see
Supplementary Material for a definition of the loglikelihoods). To evaluate the
predictive performance we average across the PLSL criteria of the
cross-validation analyses. We also assess the cross-validation variability by
means of the standard deviations of the PLSL criteria.

\section{Data Analysis}\label{DataAnalysis}

In this section we analyze the teacher data with the help of the four
cylindrical models from the previous section. We will present the results,
posterior estimates and their interpretation for all four models.

@@

\indent In the CL-GPN and GPN-SSN model we cannot compute circular regression

@@

posterior predictive distributiona are \(\hat{\theta}_{SE_{min}} = 206.87^\circ

@@

developed in the present study.\newline

\section*{References}
<div id="refs"></div>

>>>>

figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
draft             : no

@@

header-includes:
  - \DeclareRobustCommand{\VANDER}[3]{#2} 
  - \DeclareRobustCommand{\VAN}[3]{#2}
  - \DeclareRobustCommand{\DEN}[3]{#2}
  - \usepackage{multirow}
  - \usepackage{rotating}
  - \usepackage{color}
  - \usepackage{subcaption}
  - \raggedbottom
   #% set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)

@@

```{r, echo = FALSE}
library(papaja)
library(circular)
library(bpnreg)
library(haven)
library(tikzDevice)
library(plotrix)
library(tidyr)
library(dplyr) #sample_n()
library(MASS) #mvrnorm()
library(kableExtra) #latex tables
library(plotrix)
library(caret) #cross-validation
library(coda) #geweke diagnostic
```
\vspace{-0.75cm}
In social sciences the use of cylindrical data is very common. Such data consist
of a linear and a circular component. @gurtman2011reasoning refers to such data
as vectors, with a directional measure (i.e., the circular component) and a
measure indicating the magnitude (i.e., the linear component). Many established
models in psychology are often referred to as circular or circumplex models, but
those models are cylindrical. Examples of such cylindrical models are the
interpersonal circle/circumplex
[@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
affect [@russell1980circumplex], the circumplex of human emotion
[@plutchik1997general] or the model of human values [@schwartz1992values].
\newline
\indent Also, many of the more recent types of data that are studied in psychology are
cylindrical. For example, research on human navigation uses data where distance
(i.e., the linear component) and direction (i.e., the circular component) are of
interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
consist of both the direction (i.e., the circular variable) and the duration
(i.e., the linear variable) (e.g., @rayner200935th). Apart from the social
sciences, data with a circular and linear outcome more commonly occur in
meteorology [@garcia2013exploring], ecology [@garcia2014test] or
marine research [@lagona2015hidden]
\newline
\indent Up until now researchers studying cylindrical data had to rely on linear
statistical methods to analyze their research results. However, lately more and more
of these researchers acknowledge that linear methods are not sufficient and call
for new methods [@gurtman2011reasoning; @pennings2017complexity;
@wright2009integrating] that take into account both the circular and the linear
component of these data.
\newline
\indent The aim in the present paper is twofold. Firstly, we intend to fill
the above mentioned gap in the literature by showing that the use of cylindrical
models can benefit the analysis of circumplex data and cylindrical data in
psychology in general. More specifically we will show these benefits for
interpersonal teacher data from the field of educational psychology. Apart from
modelling the dependence between the linear and circular component of a
cylindrical variable we would also like to predict the two components from a set
of covariates in a regression model. Our second aim therefore is to adapt
several existing cylindrical models in such a way that they include a regression
of both the linear and circular component of a cylindrical variable onto a set
of covariates. From now on we will therefore refer to the components of the
cylindrical variable as outcome components. These adapted cylindrical models are
then used  to analyse the teacher data.

\vspace{-0.75cm}
\subsection{Cylindrical Data}
\vspace{-0.75cm}
Data that consist of a linear variable and a circular variable is called
cylindrical data. A circular variable is different from the linear variable in
the sense that it is measured on a different scale. Figure \ref{circline} shows
the difference between a circular scale (right) and a linear scale (left). The
most important difference is that on a circular scale the datapoints 0\(^\circ\)
and 360\(^\circ\) are connected and in fact represent the same number while on a
linear scale the two ends, \(-\infty\) and \(\infty\), are not connected and
consequently the values 0\(^\circ\) and 360\(^\circ\) are located on different
places on the scale. Both circular data and cylindrical data require special
analysis methods due to this periodicity in the scale of a circular variable
(see e.g. @fisher1995statistical for an introduction to circular data and
@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
elaborate overview).\newline
\indent In the literature, several methods have been put forward to model the
relation between the linear and circular component of a cylindrical variable.
Some of these are based on regressing the linear component onto the circular
component using the following type of relation: \[y = \beta_0 +
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
linear component and \(\theta\) the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
different way, e.g. by specifying a multivariate model for several linear and
circular variables and modelling their covariance matrix
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
example, @abe2017tractable introduce a cylindrical distribution based on a
Weibull distribution for the linear component and a sine-skewed von Mises
distribution for the circular component and link these through their respective
shape and concentration parameters. However, none of the methods that have been
proposed thus far include additional covariates onto which both the circular and
linear component are regressed.

@@

\vspace{-0.75cm}
\section{Four Cylindrical Regression Models}\label{Models}
\vspace{-0.75cm}
One of the goals of this paper is to show the benefits of cylindrical methods
for the analysis of circumplex data and cylindrical data in psychology in
general. To do so we decide to focus on four cylindrical models. The models were
selected for their relatively low complexity and the ease with which a
regression structure could be incorporated. But also because they show different
ways of modelling the linear and circular outcome and thereby illustrate a
wider range of cylindrical models available in the literature. The
cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_p$ for the linear and
circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
based on a construction by @mastrantonio2015bayesian, while the other models are
extensions of the models from @abe2017tractable and @mastrantonio2018joint. The
four cylindrical  models are introduced separately in the subsections below.
However, to provide a more succinct overview and comparison of the four models, Table
\ref{TableModels} gives an overview of the similarities and differences between
the models.

\begin{sidewaystable}[h]
\centering
\caption{Comparison of the four cylindrical regression models} 
\begin{tabular}{lllll}
  \noalign{\smallskip}\hline\noalign{\smallskip}
\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
$\Theta$ & &&&\\
$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
$Y$& &&&\\
$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
$\Theta$-$Y$ dependence &                                   &                                   & & \\
                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\

@@

& \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
asymmetry/skewness.\newline

@@

\begin{figure}[]
  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to. The square indicates the inflection point of the regression line.}
  \label{circregline}
\end{figure}

\vspace{-0.75cm}
\subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
\vspace{-0.75cm}
This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. It concerns a combination
of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
model defined only on the positive real half-line \([0, + \infty)\) and thus can
not be negative.\newline
\indent In this model we predict the linear scale parameter and circular
location parameter, both of which we can express in terms of
covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\gamma_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\).\newline
\indent For the teacher data, the predicted values for the circular outcome in
the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
The conditional distribution for the linear outcome is Weibull, meaning that we
can use methods from survival analsis to interpret the effect of a predictor. In
survival analysis a 'survival' function is used in which time is plotted against
the probability of survival of subjects suffering from a specific medical
condition. In the teacher data we can thus compute the probability of a teacher
having a specific strength on the IPC. This probability is computed using the
'survival-function' defined as  \[\exp(-\alpha
y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
function we also see that the circular concentration parameter \(\kappa\) and
the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
and strength of interpersonal behavior through the parameters $\beta_{0}$ and
$\gamma_0$ as well as the effect of self-efficacy on type and strength of
teacher behavior through the parameters $\beta_{1}$ and $\gamma_1$.\newline
\indent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

\vspace{-0.75cm}
\subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
\vspace{-0.75cm}
This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. The model contains \(m\)
  independent circular outcomes and \(w\) independent linear outcomes. The
  circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,
  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Because the GPN distribution is modelled
  using a so-called augmented representation (see also the description of the
  CL-PN and CL-GPN models) it is convenient to use a similar tactic for
  modelling the multivariate skew normal distribution. As in
@mastrantonio2018joint, dependence between the linear and circular outcome is
  created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
  and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
  distribution.\newline
\indent This means that we have a shared mean vector and variance-covariance
matrix for the linear and circular outcome(s), much like having multiple
outcomes in a MANOVA (multivariate analysis of variance) model. In our
regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
covariates. The mean vector then becomes \(\boldsymbol{M}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and
$\boldsymbol{x}_i$ is a $g + 1$ dimensional vector containing the value 1 to
estimate an intercept and the $g$ covariate values. This means that in contrast
to the other three models, we have to use the same set of predictors for the
circular and linear outcome.\newline
\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix}
\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} &
\beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\). The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
linear outcomes in the GPN-SSN model are
\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
\beta_{1_{s^{I}}}\text{SE}_i),\]

@@

\noindent The GPN-SSN model thus allows us to assess the average type and
strength of interpersonal behavior through the parameters $\beta_{0_{s^{I}}}$,
$\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
on type and strength of teacher behavior through the parameters
$\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
because the type and strength of interpersonal behavior are modelled together
using a multivariate normal distribution we can, through its variance-covariance
matrix, also assess the dependence between the type and strength of interpersonal
behavior. \newpage
\indent We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

\vspace{-0.75cm}
\subsection{Model Fit Criterion}\label{Modelfit}
\vspace{-0.75cm}
For the four cylindrical models we focus on their out-of-sample predictive
performance to determine the fit of the model. To do so we use k-fold
cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
of the sample) is used once as a holdout set and 9 times as part of a training
set. The analysis will thus be performed 10 times, each time on a different
training set.\newline
\indent A proper criterion to compare out-of-sample predictive performance is
the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
value of this criterion, the better the predictive performance of the model.
Using ML estimates this criterion can be computed as follows:
\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
\end{equation}
\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
Using posterior samples the criterion is similar to the log pointwise predictive
density (lppd) [@BDA, p. 169] and can be computed as:
\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
\end{equation}
\noindent where \(B\) is the amount of posterior samples and
\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
parameters for the \(j^{th}\) iteration. Because the joint density and thus also
the likelihood for the modified GPN-SSN model is not
available in closed form [@mastrantonio2018joint] we compute the PLSL for the
circular and linear outcome separately for all models. Note that although we fit
the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
prior information into account when assessing model fit with the PLSL. According
to @BDA this is not necessary since we are assessing the fit of a model
to data, the holdout set, only. They argue that the prior in such case is only
of interest for estimating the parameters of the model but not for determining
the predictive accuracy.\newline
\indent For each of the four cylindrical models and for each of the 10
cross-validation analyses we can then compute a PLSL for the circular and linear
outcome by using the conditional log-likelihoods of the respective outcome (see
Supplementary Material for a definition of the loglikelihoods). To evaluate the
predictive performance we average across the PLSL criteria of the
cross-validation analyses. We also assess the cross-validation variability by
means of the standard deviations of the PLSL criteria.

\vspace{-0.75cm}
\section{Results}\label{DataAnalysis}
\vspace{-0.75cm}
In this section we analyze the teacher data with the help of the four
cylindrical models from the previous section. We will present the results,
posterior estimates and their interpretation for all four models.

@@

\indent In the CL-GPN and GPN-SSN models we cannot compute circular regression

@@

posterior predictive distributions are \(\hat{\theta}_{SE_{min}} = 206.87^\circ

@@

developed in the present study.

\newpage
\vspace{-0.75cm}
\section*{{\normalfont References}}
\vspace{-0.75cm}
<div id="refs"></div>
